---
title: "Introduction"
author: 
- name: Jiefei Wang
  affiliation: Roswell Park Comprehensive Cancer Center, Buffalo, NY
date: "`r Sys.Date()`"
output:
    BiocStyle::html_document:
        toc: true
        toc_float: true
vignette: >
  %\VignetteIndexEntry{quickStart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
package: GCSConnection
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(GCSConnection)
gcs_cloud_auth(json_file = NULL)
```
# Credentials

## Auto authentication

You need to have a service account credentials to authenticate with
Google Cloud Storage API. The credentials determines which data
sources are accessable to you. Please follow the instructions on
[Google Authentication][] to create your credentials file and download
it to your local machine(Note that it can be done in the Cloud
Console).  The package will search for the credentials file from the
environment variable `GOOGLE_APPLICATION_CREDENTIALS` when it is
loaded. If it fails to find the credentials, it will look for the
environment variable `GCS_AUTH_FILE` instead. If both environment
variables are empty, you will only be able to access public data.  If
you set the environment variables after the package is loaded, you can
rerun the auto authentication by `gcs_cloud_auth()`.

## Manual authentication

The function `gcs_cloud_auth` provides two ways to authenticate with
Google Cloud. When you have a service account credentials, you can
authenticate with Google Cloud by calling
`gcs_cloud_auth(credentials_file_path)`. The second method is to get
credentials from [Cloud SDK][]. The package is able to interact with
the command-line tool gcloud. Once the account has been set in the
`gcloud` program, the package can be authenticated by
`gcs_cloud_auth(gcloud = TRUE)`. By default, the package will use the
first account in gcloud to authenticate with Google. If there are
multiple accounts in gcloud, you can switch the account by
`gcs_cloud_auth(gcloud = TRUE, email = "your email address")`.

## Anouymous access

Not all the cloud buckets require a credentials to access. You can
read public data without authenticating with Google. By default, if
the package fails to find the credentials from the environment
variables, it will use anouymous access method to query data from
google cloud. Otherwise, you can switch to the anouymous mode by
calling `gcs_cloud_auth(json_file = NULL)`.

Once the credentials is set, you can check your token and authenticate
type via:

```{r}
gcs_get_cloud_auth()
```

[Cloud SDK]: https://cloud.google.com/sdk/

[Google Authentication]: https://cloud.google.com/docs/authentication/production


# Manually create connections

Creating a connection to a file in a bucket is simple, you can simply
provide an URL to `gcs_connection`. Below is an example to create a
read connection to a public dataset on Google Cloud Platform.

```{r}
## equivalent:
## file <- "genomics-public-data/NA12878.chr20.sample.DeepVariant-0.7.2.vcf"
file <- "gs://genomics-public-data/NA12878.chr20.sample.DeepVariant-0.7.2.vcf"
con <- gcs_connection(description = file, open = "r")
readLines(con, n = 2L)
close(con)
```

Note that you do not need a credentials for accessing the public data.
the character "r" in the parameter `open` is an abbreviation of read,
please see `?gcs_connection` for more details.

It is also possible to specify file and bucket name separately. For
example:

```{r}
file_name <- "NA12878.chr20.sample.DeepVariant-0.7.2.vcf"
bucket_name <- "genomics-public-data"
con <- gcs_connection(
    description = file_name, open = "r", bucket = bucket_name
)
readLines(con, n = 2L)
close(con)
```

The above code can create the same connection as the previous
example. Note that you cannot provide both URI and bucket name to
`gcs_connection`.

For the write connection, it can be made by specifying an appropriate
open mode in the `gcs_connection` function. Please note that due to
the limitation of the Google Cloud Storage, the write connection is
not seekable, which means you cannot use the `seek` function to

navigate through the file. When the write connection is created.  It
always starts in the beginning of the file. If there exists a file
with the same name, the old file will be deleted. After the write
connection is closed, the file will become immutable and no further
change on the file can be made.

# create connections from the file manager

Besides manually creating a connection, the package provides a simple
and convinent S4 class to manage files. You can list all files in a
bucket/folder by

```{r}
## These are equivalent
## x <- gcs_dir("gs://genomics-public-data/clinvar/")
## x <- gcs_dir("genomics-public-data/clinvar/")

x <- gcs_dir("gs://genomics-public-data/clinvar/")
x
```

Note that for listing files in a bucket, the trailing slash does not
make any differences. However, if you want to list files in a folder
inside a bucket, it is recommended to explicitly add a `/` at the
end. The trailing slash is not mandatory, but if no trailing slash
presents, the package need one extra communication with the cloud to
determine whether the path is a file or a folder, which can double
your time cost.

In fact, there is no hierarchical structure in a bucket, all files are
in the same level. By using the slash delimiter, the cloud makes the
files appear as though they are stored in folders. Though the cloud
does not have any restriction on the use of `/` in the file name, a
good practice is to not use `/` at the end of a file name for it can
cause unnecessary confusion.

Once obtaining a list of files, you can changes your current directory
or view the detail of a file through `$` or `[[` operator

```{r}
## equivalent: x$README.txt
myfile <- x[["README.txt"]]
myfile
```

You can also use `..` to go to the parent folder, `~` to go to the bucket root.

```{r}
## equivalent: myfile$`..`
myfile[[".."]]

## myfile$`~`
myfile[["~"]]
```

The connection can be made by

```{r}
## Equivalent: gcs_connection(myfile)
con <- myfile$get_connection(open = "r")
con
close(con)
```

Note that you can pass a file object to the function `gcs_connection`. The function `gcs_cp` supports both file and folder objects. You can query file info, copy, delete files through via the file manager

```{r}
## Get file name
myfile$file_name

## copy file
## For the destination, you can specify a path to the file, 
## or a path to a folder.
destination <-tempdir()
myfile$copy_to(destination)
file.exists(file.path(destination,myfile$file_name))

## Delete file, the function is excutable 
## only when you have the right to delete the file.
## Use `quiet = TRUE` to suppress the confirmation before deleting the file.
# x$README.txt$delete(quiet = FALSE)
```

Note that you cannot delete a folder since beforementioned file
hierarchy issue. Once all files in a folder has been deleted, the
folder will be removed.

# Requester pays
Some buckets have requester pays enabled, which means you are responsible for the cost of accessing the data in the bucket. Therefore, you must have a valid billing project for reading the data in the bucket. By default, if you use JSON file to authenticate with Google Cloud, the billing project is the project listed in the JSON file, which is the project that you use to generate the JSON file. If you use gcloud, the billing project is the default project in your config file. You can view and change the default billing project via
```{r}
gcs_get_billing_project()
gcs_set_billing_project("your project ID")
```
For accessing the data in the bucket, you also need to pass the argument `billing_project = TRUE` to the gcs functions that you want to use. For example, if `uri` is a path to a file in a requester pays bucket, you can view its information via `gcs_dir(uri, billing_project = TRUE)`. Please keep in mind that you will `get charged` if you send a billing project to a bucket which do not have requester pays enabled. Therefore, the default behavior of all gcs functions is to access the bucket without a billing project. If you want your billing project being involved in every gcs function, you can alter this setting by
```{r}
gcs_set_requester_pays(TRUE)
gcs_get_requester_pays()
```


# Buffer size

For reducing the number of network requests and speeding up the
performance of the connection, there is a buffer associated with each
connection. The default size is 1Mb for a buffer. You can set or get
the buffer size via `gcs_read_buff`, `gcs_write_buff`,
`gcs_get_read_buff` and `gcs_get_write_buff` functions:

```{r}
gcs_get_read_buff()
gcs_get_write_buff()
```

Please note the minimum buffer size for a write connection is 256
Kb. Creating a connection with small buffer size may impact the
performance.


